---
title: "STAT33600 Homework 8"
author: "Sarah Adilijiang"
output:
  html_notebook: default
  pdf_document: default
---

## Section 5.1 Long Memory ARMA and Fractional Differencing

## 5.1

### (a)

```{r, warning=FALSE, fig.width=9, fig.height=4}
library(astsa)
plot(arf, main="Data Plot")
```

Comment:

There is a slowly increasing trend from time 0 to 800 and an decreasing trend from time 800 to 1000.



### (b)

```{r}
# plot of ACF & PACF
acfs = acf2(arf, max.lag=100, main="Sample ACF and PACF")

# PACF values of lag 1:5
acf(arf, type="partial", plot=FALSE)$acf[1:5]
```

Comment:

The ACF plot decays very slowly, which indicates the series is a long memory process. And the PACF plot cuts off after lag = 1, which indicates the series has the properties of an AR(1) process, i.e. $\phi(B) = 1 - \phi B$. And the value of PACF at lag = 1 is $\phi_{11} = \phi = 0.964857649$, which is close to 1 in the plot. Thus the series is like a nonstationary random walk model with $\phi(B) = 1 - B$.



### (c)

```{r, warning=FALSE}
library(fracdiff)
farima = fracdiff(arf, nar=1) # AR(1)
summary(farima)
```

Comment:

The estimated $\hat{d} = 0.264631$ and $\hat{\phi} = 0.863068$, and both are very significant.



### (d)

The data plot in (a) shows a slowly increasing trend from time 0 to 800 and an decreasing trend from time 800 to 1000.

The ACF plot in (b) tails off very slowly, and the PACF plot cuts off after lag = 1 where the value of lag = 1 is close to 1. These indicate that the data is like an AR(1) process with $\phi(B) = 1 - B$.

Therefore, it would seem reasonable to difference the data, i.e. using $\nabla x_t$, and fit an ARMA model to $\nabla x_t$.




### (e)

```{r}
x = diff(arf)
par(mfrow=c(2,1))
acfs = acf2(x, max.lag=100, main="Sample ACF and PACF of differenced data")
```

Comment:

Now both the ACF and PACF cut off after lag = 1, and the values at lag = 1 are both small. This indicates that, after the differencing, the series becomes an AR(1) process a small $\phi$, or an MA(1) process with a small $\theta$.  




### (f)

```{r}
# fit an AR(1) model to the differenced data
#sarima(diff(arf), p=1, d=0, q=0, details=FALSE)$ttable
# or fit an ARIMA(1,1,0) to the not differenced data    # same result
sarima(arf, p=1, d=1, q=0, details=FALSE)$ttable   # diffrencing: d=1

# fit an MA(1) model to the differenced data
#sarima(diff(arf), p=0, d=0, q=1, details=FALSE)$ttable
# or fit an ARIMA(0,1,1) to the not differenced data    # same result
sarima(arf, p=0, d=1, q=1, details=FALSE)$ttable   # diffrencing: d=1
```

Comment:

As mentioned in (e), the differenced series can be an AR(1) process a small $\phi$, or an MA(1) process with a small $\theta$. So I fit both these two models to the differenced data.

The AR(1) model fits the differenced data $\nabla x_t$ well with a significant small $\phi \approx 0.17$. The fitted model is:
$$(1-0.17B)\ \nabla x_t = w_t$$

The MA(1) model also fits the differenced data $\nabla x_t$ well with a significant small $\theta \approx 0.17$. The fitted model is:
$$\nabla x_t = (1+0.17B)\ w_t$$














